---
title: "Lesson1_Solution"
author: "MAJ Daniel Baller"
date: "January 4, 2022"
output:
  html_document: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Welcome to MA388 Sabermetrics! 

The purpose of today is to get to know each other and get started looking at some baseball data. If you're not proficient in R yet, no worries, please feel free to ask me or get help from someone around you.

### Data Sourcesw

There are 4 main sources of data that we will use for the course.  What info can we take from each one?

- **The Lahman Database:** 1871-2019 Basic season count statistics for for batting, pitching, and fielding.

- **Retrosheet:** Game-by-game summaries and play-by-play data of most games since 1921

- **PITCHf/x:** pitch-by-pitch data detailing the flight of every pitch since 2008.

- **Statcast via Baseball Savant:** detailed information on the position and movement of all players and the ball throughout the game since 2015.

\vspace{1in}

Let's take a look at some player data.

```{r, message = FALSE, warning = FALSE}
#Growing up my favorite player was Randy Johnson back when he was playing for the mariners.
library(Lahman)
library(tidyverse)

#get Randy Johnson's player ID
People %>%
  filter(nameLast == "Johnson", nameFirst == "Randy")
```


Let's take a look at Randy Johnson's season statistics using the Pitching table. If you don't understand the code at this point, it's okay!

\vspace{2in}

```{r, warning = FALSE}
#I want to see the opponents batting average for each team that Randy Johnson played for.

#Extract Randy Johnson's seasonal pitching data
johnson = Pitching %>% 
  filter(playerID == "johnsra05")

#calculate opponents batting average by team
johnson %>% 
  group_by(teamID) %>% 
  summarize(BAO = sum(H)/sum(BFP), Years = n())
```


```{r, fig.height=3}
#Plot his batting average by season
johnson %>% 
  ggplot(aes(x = yearID, y = BAOpp, color = teamID)) +
  geom_point() +
  labs(x = "year", 
       y = "opponents batting averagee", 
       title = "Randy Johnson's opponents batting average by year")+
  theme_classic()
```

Who is your favorite baseball player?

\vspace{0.25in}

Find your favorite player's \texttt{playerID} in the Player table below.  This ID will be used to access this player's data in other databases. See my code above for some assistance.

```{r}

```

Find out something interesting about your player using the Batting, Pitching, or Fielding tables. Feel free to use my code above.

```{r}

```

