---
title: "Lesson 4 Solution"
author: "Daniel Baller"
date: "1/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(Lahman)
library(tidyverse)
library(knitr)
```


## Last Class

Key concepts

* five tidyverse verbs

* relational data bases

Review Questions

* Using the Teams data frame, calculate average wins (W) per season for each MLB team from 2000-present.  Order your results from the best teams to the worst.

```{r}
wins2000 <- Teams %>% 
  filter(yearID >= 2000) %>% 
  group_by(teamID) %>% 
  summarize(W = sum(W),
            G = sum(G)) %>% 
  mutate(Wpct = W/G) %>% 
  arrange(-Wpct)

 wins2000 %>% 
  kable(digits = 3)
```

* Using the Teams data frame, calculate how many World Series wins (WSWin) each team had during this period and add this information to your table in the previous question.

```{r}
wins2000 <- Teams %>% 
  filter(yearID >= 2000) %>% # Filter for entries from 2000 or later
  count(WSWin, teamID) %>% # count number of wins and losses for each team
  filter(WSWin == "Y") %>% # only keep the number of wins for each team
  select(teamID, n) %>% # only select these columns
  right_join(wins2000, key = "teamID") #add to the original dataframe with a right join.  

wins2000 %>% 
  kable(digits = 3, caption = "Win percentage and World Series titles by team (2000-present).")
```

## Intro to R (Day 2)

### Split, apply, and combine data (pg 58)

It's important in data science to be able to write procedures and apply them over a data set. The Marchi text refers to this as "splitting, applying, and combining data".  In other words, we split a data frame into pieces, apply a procedure or function to each piece, and then combine the results into a new data frame.

Ways to "split, apply, and combine"...

* `group_by()` and `summarize()` functions

* `split()` and `map_df()` functions

In general, these coding tasks involve three steps:

1. write a function to perform the task on one split of the data

2. split the data on a variable

3. apply the function to each split of the data

### Example

\textit{Find the team with the most wins for each season and whether the team won the World Series.}

Two steps:

1. write a function that takes a data frame and returns the team with the most wins and whether they won the World Series

2. split the data frame on yearID

3. apply the function to each split

```{r}
#step 1 - write a function

#first write what you want the function to do on a small piece of the data
Teams %>% 
  filter(yearID==2000) %>% #Select only entries for the year 2000
  arrange(-W) %>% #arrange in decending order by number of wins
  select(teamID, W, WSWin) %>% #only select these three columns
  head(1) #pick the top one by number of wins

#next generalize function to work with whatever data you provide
mostWins <- function(data){
  data %>% 
    arrange(-W) %>% 
    select(teamID, W, WSWin) %>% 
    head(1)
}

#step 2 and 3 - split and apply
winLeaders <- Teams %>% 
  filter(yearID >= 2000) %>% #filter for entries in the year 2000 and later
  split(pull(.,yearID)) %>%  #turn the dataframe into a list where each entry is split by the yearID
  map_dfr(mostWins, .id = "yearID") #apply the function each entry of the list and provide the results as a dataframe

winLeaders %>% 
  kable(caption = "Major League win leaders by season and whether they won the World Series.")

```

\textit{Note functions are not executed until they are called.}

### Your turn

\textit{Find the team with the most home runs for each season from 2000-present.  Include the team's full name in your table instead of just the teamID.}

```{r}
#step 1 - write a function
mostHRs <- function(data){
  data %>% 
    arrange(-HR) %>% 
    select(teamID, HR, name) %>% 
    head(1)
}

#step 2 and 3 - split and apply
HRLeaders <- Teams %>% 
  filter(yearID >= 2000) %>% 
  split(pull(.,yearID)) %>% 
  map_dfr(mostHRs, .id = "yearID")

HRLeaders %>% 
  select(yearID, name, HR) %>% 
  kable(caption = "Major League win leaders by season and whether they won the World Series.")

#Difference between select and pull.  
HRLeaders %>% 
  select(yearID)

HRLeaders %>% 
  pull(yearID)
```

