---
title: "Lesson 4"
author: "Daniel Baller"
date: "1/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(Lahman)
library(tidyverse)
library(knitr)
```


## Last Class

Key concepts

* five tidyverse verbs

* relational data bases

Review Questions

* Using the Teams data frame, calculate average wins (W) per season for each MLB team from 2000-present.  Order your results from the best teams to the worst.

```{r}
wins2000 <- Teams %>% 
  filter(yearID >= 2000) %>% 
  group_by(teamID) %>% 
  summarize(W = sum(W),
            G = sum(G)) %>% 
  mutate(Wpct = W/G) %>% 
  arrange(-Wpct)

 wins2000 %>% 
  kable(digits = 3)
```

* Using the Teams data frame, calculate how many World Series wins (WSWin) each team had during this period and add this information to your table in the previous question.

```{r}

```

## Intro to R (Day 2)

### Split, apply, and combine data (pg 58)

It's important in data science to be able to write procedures and apply them over a data set. The Marchi text refers to this as "splitting, applying, and combining data".  In other words, we split a data frame into pieces, apply a procedure or function to each piece, and then combine the results into a new data frame.

Ways to "split, apply, and combine"...

* `group_by()` and `summarize()` functions

* `split()` and `map_dfr()` functions

In general, these coding tasks involve three steps:

1. write a function to perform the task on one split of the data

2. split the data on a variable

3. apply the function to each split of the data

### Example

\textit{Find the team with the most wins for each season and whether the team won the World Series.}

three steps:

1. write a function that takes a data frame and returns the team with the most wins and whether they won the World Series

2. split the data frame on yearID

3. apply the function to each split

```{r}
#step 1 - write a function



#step 2 and 3 - split and apply


```

\textit{Note functions are not executed until they are called.}

### Your turn

\textit{Find the team with the most home runs for each season from 2000-present.  Include the team's full name in your table instead of just the teamID.}

```{r}
#step 1 - write a function


#step 2 and 3 - split and apply


#Difference between select and pull.  

```

